---
title: "Project"
author: "Omima Hassan"
date: "12/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
cost = read.csv("costliv_cleaned.csv")
cost = subset(cost, select = - c(X, city))
tuition = read.csv("tuitioncost_current.csv")
tuition = subset(tuition, select = -c(Dormitory.rooms, Board3, dorms.total))
min = read.csv("min_wage_clean.csv")
min = subset(min, select = c(X2013, X2014, X2015, X2016, X2017, X2018))

tuition = tuition %>% 
  rename(Year =  ï..Year.and.control.of.institution)
```


Changing tuition to numerics instead of factors so that it's more feasible to use
```{r}
tuition$tuition2=vector("numeric", length = 1)
tuition$tuition2=(c('12349','12922','13639','14439','15505','16510','17451','18473','19364','20361','21126','22074','23011','23871','24701','25409','26132','26592','27357'))
tuition$tuition2=as.numeric(tuition$tuition2)

tuition$tuition3= vector("numeric", length=1)
tuition$tuition3= (c('7040', '7372', '7786', '8309', '9029', '9706', '10279', '10931', '11455', '12046', '12404', '12945', '13572', '14099', '14563', '14957', '15343', '15512', '15924'))
tuition$tuition3 = as.numeric(tuition$tuition3)
```



Narrowing results of states to only a few, each state is from one of the four regions.
A visual representation to how much tuition costs have risen since 1999 compared to 2018. All these values are out of state tuition rates. 

An animated plot showing the rise in tuition prices per academic year
```{r}
tuition = tuition %>% rename(Out_Of_State_Tuition = Total.tuition..fees..room..and.board)
library(plotly)
x = tuition$Year
y = tuition$Out_Of_State_Tuition

p <- tuition %>%
  plot_ly(
    x = ~Year,
    y = ~Out_Of_State_Tuition,
    frame = ~y,
    type = 'scatter',
    mode = 'markers',
    showlegend = F
    
    ) 
p
chart_link = api_create(p, filename="tuition_interactive1")
chart_link

```


After finding the trend in values for in state and out of state tuition, We wanted to see if a prediction model could predict a new value for out of state tuition based on a given value for in state tuition.
* Using the model, we predict the out of state tuition of of an academic year based on an input in state tuition
* Since the linear model is very simple, the easiest way to predict is to use the parameter estimates from the model
```{r}
require(graphics)

t1 = tuition$tuition2
t2 = tuition$tuition3


tuitionlm = lm(tuition2~tuition3, data=tuition)
summary(tuitionlm)


instate = 25300

outofstate = as.numeric((tuitionlm$coefficients["tuition3"] * instate) + tuitionlm$coefficients[1])

outofstate
```



```{r}
c1 <- cost
t1 <- tuition
min1 = min
dta = qpcR:::cbind.na(c1, t1)
dta = na.omit(dta)

dta1 = qpcR:::cbind.na(dta, min1)
dta1 = na.omit(dta1)
```


Is cost of living lower or higher for an out of state student?
Create a scatter plot showing the relationship between out of state tution and the cost of living index

Showing the cost of living out of state compared to tuition, a relationship can be seen from the plot. As cost of living is greater out of state tuition prices are lower, as tuition increases cost of living as a out of state student decreases.
```{r}
ggplot(dta, aes(x = tuition2, y = Cost.of.Living.Index)) + geom_point(stat= "identity") + labs(title = "Cost of living by Out of State Tuition", x = "Out of State Tuition", y = "Cost of Living Index")  + theme(axis.text.x = element_text(angle=90, size = 7))
```

What is the relationship between minimum wage and tuiton rates in each state in the year 2018?
Create a scatter plot visualizing minimum wage 

```{r}
a = ggplot(dta1, aes(x = state,y = tuition2)) 
a + facet_wrap(~X2018) +  geom_point() + theme(axis.text.x = element_text(angle=90, size = 7)) + labs(title = "Minimum wage by State compared with Out of State Tuition", x = "State", y = "Out of State Tuition")
```

Finding if a correlation exists between minimum wage and tuition rates for both in state and out of state students for the year 2018
Because both variables result in a negative correlation it can be said that an inverse relationship exists between tuition and minimum wage, as one value increases the other decreases and vice versa
```{r}
# Out of state and minimum wage
cor(dta1$tuition2, dta1$X2018)

# In state and minimum wage
cor(dta1$tuition3, dta1$X2018)
```









Sources:
https://www.datanovia.com/en/blog/ggplot-axis-ticks-set-and-rotate-text-labels/
https://towardsdatascience.com/create-animated-bar-charts-using-r-31d09e5841da
https://www.datanovia.com/en/lessons/rename-data-frame-columns-in-r/
https://plot.ly/r/animations/
https://stackoverflow.com/questions/3699405/how-to-cbind-or-rbind-different-lengths-vectors-without-repeating-the-elements-o
https://stats.stackexchange.com/questions/188042/how-to-predict-a-value-using-r